---
title: "Exercises 7"
author: "Your name"
date: "2024-10-07"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

# Exercises 7

# Note: remove some of the output for the PDF to be easier to read.

```{r preamble}
library(tidyverse)
```

### 1. Load the data called income.data.csv 

```{r load-data}
dat <- read_csv("income.data.csv")
```

Answer: 


### 2. What variables are included in the dataset? What stat types are they?

```{r what-variables}
names(dat)
```

Answer:



### 4. Do visual EDA for education, and gender. What is wrong with this code?

# Once you do it correctly, what can you learn from the plot?

```{r visual-EDA-cat}

dat %>% ggplot(aes(x=education, y=gender)) + geom_point()


```

Answer: 


### 5. Do quantitative EDA for education and gender together.

```{r quant-EDA-cat}

```


Answer: 


### 6. Do visual EDA on income and happiness. What can you learn from this? (Remember to note all the features of the plot.)

```{r visual-EDA-quant}

```

Answer:





### 7. Do quantitative EDA for happiness and income. Comment on what you see.

```{r quant-EDA-quant}

```

Answer: 




### 8.a. Draw a scatterplot of happiness vs. income that has different colors by gender.

### 8.b. Draw a scatterplot of happiness vs. income that has different panels by gender, using facet_grid.

# What is the difference between them? What can you learn from them? Which one do you think is clearer?

```{r}

```

Answer: 

