---
title: "Practice for Exam 1"
author: "Your name"
date: "2024-09-25"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Practice for Exam 1

```{r preamble}

```

### 1. Load the data called police_use_of_force. Data on: Minneapolis Police Stops and Police Violence. Source: https://www.kaggle.com/datasets/paultimothymooney/minneapolis-police-stops-and-police-violence/

```{r load-data}

```

Answer: 


### 2. What variables are included in the dataset?

```{r what-variables}

```

Answer:


### 3. What type of stat variables are ForceType, Race, and EventAge and what values can they take? Comment on anything unusual you find.

```{r ForceType-Race-EventAge-type}

```

Answer:


### 4. Do visual EDA for ForceType and Race. Do you find anything interesting or unusual?

```{r ForceType-Race-visual-EDA}

```

Answer: 


### 5. Do quantitative EDA for ForceType and Race. Which do you think is more helpful, a plot or a table?

```{r ForceType-Race-quant-EDA}

```

Answer: 


### 6. Do visual EDA on EventAge. What can you learn from this?

```{r EventAge-visual-EDA}

```

Answer:


### 7. Do quantitative EDA for EventAge. Comment on what you see.

```{r EventAge-quant-EDA}

```

Answer: 

