---
title: "Exercises 5"
author: "Maria Cuellar"
date: "2024-09-20"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercises 5

```{r preamble}
#install.packages("tidyverse")
library(tidyverse)
```

### 1. Load the data called domestic violence

```{r load-data}
dat <- read_csv("data/domestic_violence.csv")
dat

```

Answer: I loaded the data.

### 2. What type of stat variable is Employment?

```{r Employment-type}

```

Answer: 


### 3. What categories does it have?

```{r Employment-categories}

```

Answer: 


### 4. Do quantitative EDA for Employment.

```{r Employment-quant-EDA}

```

Answer: 


### 5. Do visual EDA for Employment.

```{r Employment-visual-EDA}

```

Answer: 

### 6. What kind of variables are `Marital status` (and why does it have single quotes around it) and violence?
Categorical, and `Marital status` has quotes because it has a space in the name.


### 7. Make a contingency table of both Marital status and Violence
```{r contingency-table}

```

Answer: 

