---
title: "Crim 1200 - Statistics for the Social Sciences"
author: "Written by Maria Cuellar - updated on "
date: "2025-09-15"
format: 
  gfm:
    toc: false
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 2

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-data, echo=FALSE, include=FALSE}

library(tidyr)
library(readr)
library(dplyr)
library(tidyverse)
library(usmap)
library(readxl)
```


Instructor website: https://web.sas.upenn.edu/mcuellar/

## Course description:

Statistics is the science of developing and studying methods for collecting, analyzing, interpreting, and
presenting empirical data. A common challenge in statistics is dealing with incomplete information: when
the entire picture is not visible, how can findings from a small sample be generalized to a broader
population, and should they be?

This course offers an introduction to statistics with a focus on careful and systematic data analysis. Students
will learn to critically assess how datasets are collected, approach them thoughtfully, and decide on the
appropriate analyses to answer specific research questions. Rather than simply following a set procedure,
students will gain the tools to determine which methods are best suited for the data—and understand the
implications when ideal procedures aren’t feasible. They will also learn to identify the assumptions needed
to make credible claims and draw valid conclusions.

Designed as a foundational course in statistics for the social sciences, this class is suitable for students
pursuing a bachelor’s degree in criminology or related fields. It requires minimal mathematical or statistical
background. By the end of the course, students should be able to explore datasets using visualization
techniques, select and perform the appropriate statistical analyses for research, accurately interpret the
results of analyses and effectively communicate them to others and identify when to seek additional help
with designs or analyses beyond the scope of this course.

## Workshops:

- [Downloading R (up to section 1.2.1)](https://moderndive.netlify.app/1-getting-started.html)
- [First look at the data](https://mariacuellar.github.io/crim_data_analysis/workshops/firstlook.html)
- [ggplot](https://mariacuellar.github.io/crim_data_analysis/workshops/ggplot.html)
- [Exploring distribution](https://github.com/mariacuellar/crim_data_analysis/blob/main/exercises/Exercises%20%234.R)


## Exercises:

1. [Reading data in RStudio](https://github.com/mariacuellar/crim_data_analysis/blob/main/exercises/Exercises%201%20-%20questions.R)
[(Solutions)](https://github.com/mariacuellar/crim_data_analysis/blob/main/exercises/Exercises%201%20-%20solutions.R)
2. [First look at the data](https://github.com/mariacuellar/crim_data_analysis/blob/main/exercises/Exercises%202%20-%20questions.R)
(Solutions)
3. [Visual summary of quantitative and categorical data](https://github.com/mariacuellar/crim_data_analysis/blob/main/exercises/Exercises%203%20-%20questions.R)
(Solutions)

## Datasets:

To download the data, right-click the link and choose “Save Link As…”. Save the data into the folder of your choice.

1. [Kentucky derby](data/kentucky-derby-2018.csv)
2. [Students](data/students.csv)
3. [Domestic violence](data/domestic_violence.csv)

