---
title: "Crim 1200 - Statistics for the Social Sciences"
author: "Written by Maria Cuellar"
date: "2024-12-11"
format: 
  gfm:
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-data, echo=FALSE, include=FALSE}

library(tidyr)
library(readr)
library(dplyr)
library(tidyverse)
library(usmap)
library(readxl)
```


Instructor website: https://web.sas.upenn.edu/mcuellar/

## Course description:

Statistics is the science of developing and studying methods for collecting, analyzing, interpreting, and
presenting empirical data. A common challenge in statistics is dealing with incomplete information: when
the entire picture is not visible, how can findings from a small sample be generalized to a broader
population, and should they be?

This course offers an introduction to statistics with a focus on careful and systematic data analysis. Students
will learn to critically assess how datasets are collected, approach them thoughtfully, and decide on the
appropriate analyses to answer specific research questions. Rather than simply following a set procedure,
students will gain the tools to determine which methods are best suited for the data—and understand the
implications when ideal procedures aren’t feasible. They will also learn to identify the assumptions needed
to make credible claims and draw valid conclusions.

Designed as a foundational course in statistics for the social sciences, this class is suitable for students
pursuing a bachelor’s degree in criminology or related fields. It requires minimal mathematical or statistical
background. By the end of the course, students should be able to explore datasets using visualization
techniques, select and perform the appropriate statistical analyses for research, accurately interpret the
results of analyses and effectively communicate them to others and identify when to seek additional help
with designs or analyses beyond the scope of this course.

## Flowchart:

Here is a flowchart to do exploratory data analysis, or EDA. 

![Flowchart](../graphics/flowchart.pdf){width=65%}

## Exercises:


1. [Reading data in RStudio](https://github.com/mariacuellar/crim_data_analysis/blob/main/exercises/Exercises%20%231.R)
2. [First look at the data](https://github.com/mariacuellar/crim_data_analysis/blob/main/exercises/Exercises%20%232.R)
3. [EDA for a single categorical variable](https://github.com/mariacuellar/crim_data_analysis/blob/main/exercises/Exercises%20%233.R)
4. [EDA for a single quantitative variable](https://github.com/mariacuellar/crim_data_analysis/blob/main/exercises/Exercises%20%234%20with%20notes.R)
5. [EDA for two categorical variables](http://htmlpreview.github.io/?https://github.com/mariacuellar/crim_data_analysis/blob/main/exercises/Exercises--5---with-notes.html)
6. [Practice for exam 1](http://htmlpreview.github.io/?https://github.com/mariacuellar/crim_data_analysis/blob/main/exercises/Exercises--6.html)
7. [EDA for a categorical and quantitative variable variable](http://htmlpreview.github.io/?https://github.com/mariacuellar/crim_data_analysis/blob/main/exercises/Exercises--7-with-notes.html)
8. [Diagnostics of a linear model](http://htmlpreview.github.io/?https://github.com/mariacuellar/crim_data_analysis/blob/main/exercises/Exercises--8-with-notes.html)
9. [Preparation for exam - Linear models](http://htmlpreview.github.io/?https://github.com/mariacuellar/crim_data_analysis/blob/main/exercises/Exercises--9.html)


## Datasets:

1. [Kentucky derby data](../main/data/kentucky-derby-2018.csv)
2. [Crime in Vancouver data](../main/data/crimeinvancouver.csv)
3. [Domestic violence data](../main/data/domestic_violence.csv)
4. [Police use of force](../main/data/police_use_of_force.csv)
5. [Income data](../main/data/income.data.csv)
6. [Penguin dives](../main/data/penguin_dives.csv)
7. [Crime rate and economic stability](../main/data/crimerate_econstability.csv)



Inspiration from : https://github.com/gregridgeway/R4crim/blob/master/README.md?plain=1