---
title: "Crim 1200 Statistics for the Social Sciences"
format:
  gfm: default     # Always produce a .md file for GitHub
  html:
    toc: true
    toc-depth: 2
    number-sections: false
    toc-location: left
    favicon: upenn-favicon.ico

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-data, echo=FALSE, include=FALSE}

library(tidyr)
library(readr)
library(dplyr)
library(tidyverse)
library(usmap)
library(readxl)
```

Instructor website: https://web.sas.upenn.edu/mcuellar/

## Course description

Statistics is the science of developing and studying methods for
collecting, analyzing, interpreting, and presenting empirical data. A
common challenge in statistics is dealing with incomplete information:
when the entire picture is not visible, how can findings from a small
sample be generalized to a broader population, and should they be?

This course offers an introduction to statistics with a focus on careful
and systematic data analysis. Students will learn to critically assess
how datasets are collected, approach them thoughtfully, and decide on
the appropriate analyses to answer specific research questions. Rather
than simply following a set procedure, students will gain the tools to
determine which methods are best suited for the data—and understand the
implications when ideal procedures aren’t feasible. They will also learn
to identify the assumptions needed to make credible claims and draw
valid conclusions.

Designed as a foundational course in statistics for the social sciences,
this class is suitable for students pursuing a bachelor’s degree in
criminology or related fields. It requires minimal mathematical or
statistical background. By the end of the course, students should be
able to explore datasets using visualization techniques, select and
perform the appropriate statistical analyses for research, accurately
interpret the results of analyses and effectively communicate them to
others and identify when to seek additional help with designs or
analyses beyond the scope of this course.

## Workshops

-   [Downloading R (up to section
    1.2.1)](https://moderndive.netlify.app/1-getting-started.html)
-   [First look at the
    data](https://mariacuellar.github.io/crim_data_analysis/workshops/firstlook.html)
-   [EDA](https://mariacuellar.github.io/crim_data_analysis/workshops/EDA.html)
-   [R Markdown](https://mariacuellar.github.io/crim_data_analysis/workshops/Rmarkdown.html)
-   [Linear
    regression](https://mariacuellar.github.io/crim_data_analysis/workshops/LinearRegression.html)



## Exercises

-  [Exercises 1: Reading data in RStudio](https://raw.githubusercontent.com/mariacuellar/crim_data_analysis/refs/heads/main/exercises/Exercises%201%20-%20questions.R)
    [(Solutions)](https://raw.githubusercontent.com/mariacuellar/crim_data_analysis/refs/heads/main/exercises/Exercises%201%20-%20solutions.R)
-  [Exercises 2: First look at the data](https://raw.githubusercontent.com/mariacuellar/crim_data_analysis/refs/heads/main/exercises/Exercises%202%20-%20questions.R)
    [(Solutions)](https://raw.githubusercontent.com/mariacuellar/crim_data_analysis/refs/heads/main/exercises/Exercises%202%20-%20solutions.R)
-  [Exercises 3: Visual summary of quantitative and categorical data](https://raw.githubusercontent.com/mariacuellar/crim_data_analysis/refs/heads/main/exercises/Exercises%203%20-%20questions.R) 
[(Solutions)](https://raw.githubusercontent.com/mariacuellar/crim_data_analysis/refs/heads/main/exercises/Exercises%203%20-%20solutions.R)
-  [Exercises 4: Using R Markdown](https://raw.githubusercontent.com/mariacuellar/crim_data_analysis/refs/heads/main/exercises/Exercises%204%20-%20questions.R) 
[(Rmd Solutions)](https://raw.githubusercontent.com/mariacuellar/crim_data_analysis/refs/heads/main/exercises/Exercises%204%20-%20solutions.rmd)
[(HTML Solutions)](https://mariacuellar.github.io/crim_data_analysis/exercises/Exercises-4---solutions.html
)
- [Practice exercises for Exam 1 - in RMarkdown](https://raw.githubusercontent.com/mariacuellar/crim_data_analysis/refs/heads/main/exercises/Exam 1 practice - questions.Rmd) 
[(Solutions)](https://raw.githubusercontent.com/mariacuellar/crim_data_analysis/refs/heads/main/exercises/Exam%201%20practice%20-%20solutions.Rmd)




## Datasets

To download the data, right-click the link and choose “Save Link As…”.
Save the data into the folder of your choice.

1.  [kentucky-derby-2018.csv](data/kentucky-derby-2018.csv)
2.  [students.csv](data/students.csv)
3.  [domestic_violence.csv](data/domestic_violence.csv)
4.  [pretrial_df](data/pretrial_df.csv) [(Codebook for pretrial_df)](https://raw.githubusercontent.com/mariacuellar/crim_data_analysis/refs/heads/main/data/codebook%20for%20pretrial_df.txt)

