---
title: "Exam 1 Questions"
author: "Your name"
date: "2026-09-28"
output:
  html_document: default
  pdf_document: default
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



*Note: For this exam, don't worry about including the correct labels and titles for the plots.*

```{r preamble, echo=FALSE, message=FALSE}
library(tidyverse)
```

(30 possible points)

### Question 1.

Load the data called riverton-crime.csv using the proper R command and file path. What is this dataset about, according to the codebook? (Keep it short.)

**Your code (2 points):** 

```{r load-data, message=FALSE}
2+2
```

**Your text answer (2 points):**






---

### Question 2.

What types of variables are these? Variables: id, gender, offense_type, num_prior_arrests, and time_served_months.

**Your text answer (2 points):**

- id: 
- gender: 
- offense_type: 
- num_prior_arrests: 
- time_served_months: 



---

### Question 3.

What percentage of the sample is male? How many females, males, and total individuals are there in this sample?



**Your code (2 points):**

```{r male}
2+2
```



**Your text answer (2 points):**







---

### Question 4. 

Draw a plot to study the variable offense_type and describe it.



**Your code (2 points):**

```{r offense_type}
2+2
```



**Your text answer (2 points):**






---

### Question 5.

Now we are going to study the variable num_prior_arrests. 

Perform visual EDA on num_prior_arrests. Adjust the number of bins so that the shape of the distribution is reasonably clear.



**Your code (2 points):**

```{r num_prior_arrests, fig.width=3, fig.height=3}
2+2
```


---

### Question 6.

Describe this plot.


**Your text answer (2 points):**





---

### Question 7.

Find summary statistics to describe the centrality and spread of the variable num_prior_arrests. Which is the centrality and which is the spread? Explain what these values mean for this variable.



**Your code (2 points):**

```{r num_prior_arrests_stats}
2+2
```


**Your text answer (2 points):**






---

### Question 8.

Now we will study the variable time_served_months. What is this variable measuring? Visualize this variable by drawing a plot.

**Your code (2 points):**

```{r time_served_months}
2+2
```

**Your text answer (2 points):**






---

### Question 9.

Describe the shape of the distribution. Then perform quantitative EDA on it.

**Your code (2 points):**

```{r time_served_months-quant}
2+2
```

**Your text answer (2 points):**





---

### Extra credit:

Describe what this code does. Why do the points appear in vertical columns rather than being spread continuously across the x-axis?

**Your code (1 point):**

```{r scatterplot}
2+2
```

**Your text answer (1 point):**

