---
title: "First look at the data"
author: "Maria Cuellar"
date: "2025-09-16"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




# Change appearance colors in RStudio

Navigate to Tools > Global Options > Appearance (or RStudio > Preferences > Appearance on macOS) and select a desired theme from the dropdown menu. 



# Load data

Load the tidyverse - a package that will give you lots of tools.

```{r tidyverse}
library(tidyverse)
```

Load data (make sure you name it as an object with a useful name)

```{r load-data}
dat_kentucky <- read_csv("/Users/mariacuellar/Github/crim_data_analysis/data/kentucky-derby-2018.csv")
dat_students <- read_csv("/Users/mariacuellar/Github/crim_data_analysis/data/students.csv")
```




# Data selection

How to select an observation: 

```{r}
year1986 <- dat_kentucky %>% filter(Year==1986)

year1986
```



Note: understand the pipe operator (%>% or |>): The main function of the pipe operator is to take the output of the expression or function on its left-hand side (LHS) and pass it as the first argument to the function on its right-hand side (RHS). This allows you to chain multiple operations together in a clear, sequential manner.


# How to select a variable:

```{r}
dat_kentucky$Winner

thewinner <- dat_kentucky %>% select(Winner)

thewinner
```

How to select both (observation and variable):

```{r}
thewinnerin1986 <- dat_kentucky %>% 
  filter(Year==1986) %>% 
  select(Winner)

thewinnerin1986
```



# First look at the data

Variables in dataset? can look at Data pane, or use names()
```{r}
names(dat_kentucky)
```

Type of R variable: can look at the Data pane, or use class()
```{r}
class(dat_kentucky$Secs)
```

Look at dimensions of data: can look at the Data pane, or use dim()
```{r}
dim(dat_kentucky)
```

*Note*: the commands let me see what you did.




# Summarize categorical variable

How to make a table to summarize a categorical variable:

```{r}
names(dat_students)

table(dat_students$year_in_college) # Base R

dat_students %>% count(year_in_college) # tidyverse
```

How to make add proportions:

```{r}
dat_students %>% 
  count(year_in_college) %>% 
  mutate(prop = n / sum(n))
```

How to make a table to summarize Two categorical variables:
```{r}
table(dat_students$favorite_color, dat_students$year_in_college) # It's trickier in tidyverse, but check out the janitor package
```





# Visualize quantitative variable

How to draw a histogram:

```{r}
dat_kentucky %>% ggplot(aes(x=Time.in.Sec)) + geom_histogram()

dat_kentucky %>% ggplot(aes(x=Time.in.Sec)) + geom_histogram(bins=60) # Changes number of bins
```



```{r}
dat_kentucky_lowtime <- dat_kentucky %>% filter(Time.in.Sec<140)

dat_kentucky_lowtime %>% ggplot(aes(x=Time.in.Sec)) + geom_histogram()



dat_kentucky %>% 
  filter(Time.in.Sec < 140) %>% 
  ggplot(aes(x=Time.in.Sec)) + 
  geom_histogram()

```


You can save the plot as an object 

```{r}
p <- dat_kentucky %>% ggplot(aes(x=Time.in.Sec)) + geom_histogram()

library(ggthemes)
p + theme_economist()
p + labs(title="Histogram") + theme_economist()
```


How to change the aesthetics and labels

```{r}
dat_kentucky %>% ggplot(aes(x=Time.in.Sec)) + geom_histogram(bins = 30, fill = "steelblue", color = "white")
```
  

How to change the aesthetics and labels
```{r}
dat_kentucky %>% ggplot(aes(x=Time.in.Sec)) + geom_histogram() +
    labs(
    x = "Time in seconds",   # new label for x-axis
    y = "Count",         # new label for y-axis
    title = "Histogram of Time in Seconds"
  )
```

This is equivalent to
```{r}
p +
  labs(
    x = "Time in seconds",   # new label for x-axis
    y = "Count",         # new label for y-axis
    title = "Histogram of Time in Seconds"
  )
```

You can change the theme too
```{r}
p + theme_economist()
p + theme_stata()
p + theme_fivethirtyeight()
p + theme_minimal()
```




# Visualize categorical variable

How to draw a barplot for YearInCollege

```{r}
# using base R, it's just barplot()
barplot(table(dat_students$year_in_college))
```

```{r}
dat_students %>% ggplot(aes(x = year_in_college)) + geom_bar()
```

