---
title: "Linear regression"
author: "Maria Cuellar"
output:
  html_document:
    toc: false
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 4.5)
```

# Getting started

We will use simple linear regression to describe the association between two quantitative variables, interpret a fitted line, and make predictions. All data below are **fictional teaching examples**, not evidence about real courts or communities. Only base R is needed.

[Download the R Markdown source](linear-regression-1.Rmd). Open it in RStudio and run the code chunks in order, or click **Knit** to create an HTML report. Knitting requires the `rmarkdown` package; the analyses themselves use only base R. The code is also displayed throughout this page. Each row in these examples represents a different observational unit.

# Example 1: Court caseload and waiting time

Do courts with larger daily caseloads tend to have longer average waiting times? The explanatory variable, x, is daily caseload (cases); the response, y, is average waiting time (minutes).

```{r court-data}
courts <- data.frame(
  caseload = c(12, 15, 18, 20, 22, 25, 28, 30, 32, 35,
               38, 40, 42, 45, 48, 50, 52, 55, 58, 60),
  wait_minutes = c(23, 31, 26, 35, 29, 40, 38, 35, 46, 41,
                   49, 43, 54, 48, 59, 53, 63, 57, 68, 62)
)
head(courts)
summary(courts)
```

## Plot the data

Always look at the relationship before fitting a line. Look for curvature, unusual observations, and whether the vertical spread changes with x.

```{r court-scatterplot}
plot(wait_minutes ~ caseload, data = courts,
     pch = 19, col = "steelblue",
     xlab = "Daily caseload (cases)",
     ylab = "Average waiting time (minutes)",
     main = "Caseload and waiting time in fictional courts")
```

## Fit and inspect the model

In R, the formula `y ~ x` means model y using x. `lm()` fits the line that minimizes the sum of squared residuals.

```{r court-model}
model <- lm(wait_minutes ~ caseload, data = courts)
summary(model)
coef(model)
confint(model)

library(tidyverse)
courts %>% ggplot(aes(x=caseload, y=wait_minutes)) + geom_point() + geom_smooth(method="lm", se=FALSE)

```

The fitted equation is: predicted waiting time = intercept + slope × caseload.

- **Intercept:** predicted waiting time when caseload is zero. Zero is outside our observed range, so this is not a useful substantive prediction here.
- **Slope:** the estimated difference in average waiting time, in minutes, associated with one additional daily case.
- **Slope p-value:** tests the null hypothesis that the population slope is zero, under the model assumptions.
- **95% confidence interval:** gives a range of plausible values for the population slope under those assumptions.
- **R-squared:** the proportion of sample variation in waiting time explained by the fitted line. It does not establish causation.

```{r court-interpretation}
# Extract the quantities used in a written interpretation.
slope <- unname(coef(model)["caseload"])
slope_ci <- confint(model)["caseload", ]
r_squared <- summary(model)$r.squared
round(c(slope = slope, lower_95 = slope_ci[1],
        upper_95 = slope_ci[2], R_squared = r_squared), 3)
```

Here, one additional daily case is associated with an estimated **`r round(slope, 2)` extra minutes** of average waiting time (95% CI: `r round(slope_ci[1], 2)` to `r round(slope_ci[2], 2)`). Courts were not randomly assigned caseloads: staffing, case complexity, or other factors could help explain this association.

## Make predictions

Predict within the observed range. A confidence interval describes uncertainty about the mean response at a given x. A prediction interval describes uncertainty about the response for one new court at that x, so it is wider.

```{r court-predictions}
range(courts$caseload)
new_courts <- data.frame(caseload = c(25, 45))
predict(model, newdata = new_courts, interval = "confidence")
predict(model, newdata = new_courts, interval = "prediction")
```

## Check residuals and assumptions

For conventional confidence intervals and tests, we assume a linear mean relationship, independent errors, constant error variance, and approximately normal errors (especially for small samples). Independence comes from the study design; a plot cannot establish it. Repeated observations from the same court would require additional care.

```{r court-diagnostics}
# A residual is observed minus fitted waiting time.
head(data.frame(observed = courts$wait_minutes,
                fitted = fitted(model), residual = resid(model)))

# Ideally, residuals form a roughly even cloud around zero.
plot(fitted(model), resid(model), pch = 19,
     xlab = "Fitted waiting time (minutes)", ylab = "Residual (minutes)")
abline(h = 0, lty = 2)

# Points should be approximately along the line.
qqnorm(resid(model))
qqline(resid(model), col = "firebrick")

# Large Cook's distances flag potentially influential courts to investigate. Can make the Residuals vs Leverage plot or just look at the Cook's distances directly.
plot(model, which = 5)

plot(model, which = 4)
```

Curvature suggests that a straight line may be inadequate; a funnel shape suggests unequal variance. Investigate unusual observations rather than automatically deleting them. These checks are limited with small samples.



```{r}

courts_dat_clean <- courts %>%
  filter(cooks.distance(model) <= 0.05)

```



# Example 2: Neighborhood resources and reported incidents

This second fictional example has a negative association. Each row represents a neighborhood. The outcome is a reported incident rate per 1,000 residents, not an individual person's risk.

```{r neighborhood-example}
neighborhoods <- data.frame(
  recreation_hours = c(2, 4, 5, 7, 8, 10, 11, 13, 14, 16, 18, 20),
  incidents_per_1000 = c(38, 36, 39, 32, 35, 30, 33, 27, 29, 24, 27, 21)
)
model2 <- lm(incidents_per_1000 ~ recreation_hours, data = neighborhoods)
plot(incidents_per_1000 ~ recreation_hours, data = neighborhoods,
     pch = 19, col = "steelblue",
     xlab = "Weekly recreation program hours",
     ylab = "Reported incidents per 1,000 residents")
abline(model2, col = "firebrick", lwd = 2)
summary(model2)
confint(model2)
predict(model2, newdata = data.frame(recreation_hours = 12),
        interval = "prediction")
```

Interpret the slope as the estimated difference in reported incidents per 1,000 residents associated with one additional weekly program hour. An association between neighborhoods does not establish that adding program hours causes a change in crime or that the same relationship holds for individuals. Check the assumptions as in Example 1 before relying on inference.

# Try it yourself

1. Use the first model to predict waiting time for a court with 35 daily cases. Compare its confidence and prediction intervals.
2. Explain why predicting a court with 200 daily cases would be risky.
3. Interpret the slope and its confidence interval in Example 2, including units.
4. Give one potential confounder in each example.

# Reporting a regression

State the observational unit, outcome and predictor, slope with units and a confidence interval, and what the diagnostics suggest. Explain the limits of the design. Statistical significance does not measure the size or practical importance of an association.
