---
title: "Linear regression"
author: "Maria Cuellar"
output:
  html_document:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 4.5)
```



# Example 1: Court caseload and waiting time

These are fictional courts. **x** is daily caseload; **y** is mean waiting time in minutes. Court 21 is deliberately unusual.

## 1. Explore the data

```{r eda}
library(tidyverse)
courts <- data.frame(
  court_id = 1:21,
  caseload = c(12, 15, 18, 20, 22, 25, 28, 30, 32, 35,
               38, 40, 42, 45, 48, 50, 52, 55, 58, 60, 70),
  wait_minutes = c(23, 31, 26, 35, 29, 40, 38, 35, 46, 41,
                   49, 43, 54, 48, 59, 53, 63, 57, 68, 62, 20)
)

courts %>%
  ggplot(aes(x = caseload, y = wait_minutes)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Daily caseload", y = "Mean waiting time (minutes)")

cor(courts$caseload, courts$wait_minutes)
```

Describe direction, strength, shape, and unusual points. Correlation summarizes a **linear** association and is sensitive to outliers.

## 2. Fit the full model

```{r full-model}
model_full <- lm(wait_minutes ~ caseload, data = courts)
summary(model_full)
```

In the coefficient table, use the `caseload` row for the slope and its p-value.

## 3. Check influence and refit if justified

```{r influence}
cook <- cooks.distance(model_full)
data.frame(court_id = courts$court_id,
           cooks_distance = round(cook, 3))
plot(model_full, which = 4)
```

Court 21 has a large Cook's distance. We omit it **only for a sensitivity analysis** and retain the original model.

```{r refit}
courts_without_21 <- filter(courts, court_id != 21)
model_without_21 <- lm(wait_minutes ~ caseload, data = courts_without_21)
summary(model_without_21)
```

When an exercise specifies a Cook's distance cutoff (such as 0.05), this is the filtering pattern. It assumes the fitted model used every row in the same order; check for missing values and inspect the flagged rows before excluding them.

```{r cutoff-pattern, eval=FALSE}
model <- lm(y ~ x, data = dat)
dat_nooutliers <- dat %>% filter(cooks.distance(model) <= 0.05)
model_nooutliers <- lm(y ~ x, data = dat_nooutliers)
```

## 4. Compare both models

```{r comparison}
# Correlation can change when an influential point is removed.
cor(courts$caseload, courts$wait_minutes)
cor(courts_without_21$caseload, courts_without_21$wait_minutes)

# Compare the slope, its p-value, and R-squared in each summary.
summary(model_full)
summary(model_without_21)

# R's four standard diagnostic plots for each model.
par(mfrow = c(2, 2))
plot(model_full)
plot(model_without_21)
par(mfrow = c(1, 1))
```

| Check | What to look for |
|---|---|
| Linearity | Roughly straight scatterplot; no clear curve in residuals |
| Constant variance | Similar residual spread across fitted values |
| Normality | Q-Q points roughly follow the line |
| Independence | Does the study design make observations independent? |

Be reasonable with real data: small departures are common. The residuals vs leverage plot also helps identify influential points. Report if the conclusion changes between models.

## 5. Interpret

Use this sentence: **For one additional case per day, predicted mean waiting time changes by [slope] minutes, on average.** Then state whether the slope is statistically significant at 0.05, give its p-value, and describe association strength using correlation or R-squared. Read these from `summary(model_full)` and `summary(model_without_21)`.

Association does not establish that adding cases *causes* waiting time to change.

# Example 2: Try the workflow

Each row is a fictional neighborhood. **x** is weekly recreation program hours; **y** is reported incidents per 1,000 residents.

```{r second-example}
neighborhoods <- data.frame(
  recreation_hours = c(2, 4, 5, 7, 8, 10, 11, 13, 14, 16, 18, 20),
  incidents_per_1000 = c(38, 36, 39, 32, 35, 30, 33, 27, 29, 24, 27, 21)
)
neighborhoods %>%
  ggplot(aes(x = recreation_hours, y = incidents_per_1000)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
cor(neighborhoods$recreation_hours, neighborhoods$incidents_per_1000)
model2 <- lm(incidents_per_1000 ~ recreation_hours, data = neighborhoods)
summary(model2)
par(mfrow = c(2, 2))
plot(model2)
par(mfrow = c(1, 1))
```

Describe the slope in **incidents per 1,000 residents per additional program hour**. Give the p-value and say whether the association is strong. Can you interpret it as causal? Why or why not?
