---
title: "RMarkdown"
author: "Maria Cuellar"
date: "2025-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

R Markdown lets you **combine code, output, and text** in the same document.
You can write your analysis, include plots, and explain your results — all in
one file — and render it into an HTML, PDF, or Word report.

This guide will help you:

1. Understand the basic structure of an `.Rmd` file  
2. Write text using **Markdown** formatting  
3. Add and run R code chunks  
4. Render the document to HTML  

---

# R Markdown Structure

An R Markdown file has three parts:

1. **YAML header** – appears at the top between `---` lines (title, author, etc.)
2. **Markdown text** – regular writing with formatting
3. **Code chunks** – blocks of R code surrounded by three backticks

---

# Writing Text with Markdown

You can write *italic*, **bold**, and even lists:

- Item 1  
- Item 2  
- Item 3  

You can also add links:  
[Visit RStudio's R Markdown site](https://rmarkdown.rstudio.com)

And equations (if you want):  
`$E = mc^2$` → \(E = mc^2\)

---

# Code Chunks

To include R code, use a **code chunk** using this syntax: <pre> ```{r}  x <- 5 x + 10 ``` </pre>

Ends up looking like this:

```{r}
# A simple calculation
x <- 5
y <- 10
x + y
```



# Adding Plots

Plots can be added using the same 'chunk' syntax:

```{r}
# Make a simple scatterplot

plot(mtcars$mpg, mtcars$wt,
     xlab = "Miles per Gallon",
     ylab = "Weight",
     main = "Car Weight vs MPG")
```



# Chunk Options

You can control what appears in your report with chunk options in with the {r} curly braces: e.g., <pre> ```{r, echo=FALSE, eval=FALSE}  x <- 5 x + 10 ``` </pre>:

- echo=FALSE → hide code, show results
- eval=FALSE → show code, but don’t run it
- message=FALSE, warning=FALSE → hide messages/warnings



# Rendering Your Document

Click Knit in RStudio and choose Knit to HTML.

R will:

- Run all the code chunks
- Combine them with your text
- Produce a clean, shareable report

# Tips for Success:

- Save your .Rmd file often.
- Knit regularly to catch errors early.
- Use clear headings (#, ##, ###) to organize your report.
- Keep code readable: use comments (#) to explain what it does.


# Practice

Try using the mtcars dataset and adding:

1. A summary statistic, mean of disp
2. A histogram of disp
3. A short paragraph explaining what you found
4. Then knit again and see how your document updates!

