---
title: "Linear Regression In-Class Exercises"
author: "Maria Cuellar"
date: "2026-10-05"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)

dat <- read_csv("/Users/mariacuellar/Github/crim_data_analysis/data/riverton-crime.csv")
```

# Linear Regression

We will examine the relationship between the number of prior arrests and time served.

## Question 1

Create a scatterplot with `num_prior_arrests` on the x-axis and `time_served_months` on the y-axis.

```{r}
dat %>%
  ggplot(aes(x = num_prior_arrests, y = time_served_months)) +
  geom_point()
```

Describe the relationship. Is it positive or negative? Does it look approximately linear?

**Answer:**

Positive, kind of linear but maybe a little curved down.

## Question 2

Fit a linear regression where you regress time served onto number of prior arrests.

```{r}
model <- lm(time_served_months ~ num_prior_arrests, data = dat)

summary(model)
```

Which variable is the outcome/dependent variable?

**Answer:**

Which variable is the explanatory/independent variable?

**Answer:**


## Question 3

Use the regression output to write the regression equation.

Predicted time served = ______ + ______ × number of prior arrests

Interpret the slope in words.

**Answer:**


## Question 4

Add the regression line to the scatterplot.

```{r}
dat %>%
  ggplot(aes(x = num_prior_arrests, y = time_served_months)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

Does the regression line seem like a reasonable summary of the relationship?

**Answer:**


## Question 5

Use the model to predict time served for a person with 5 prior arrests.

```{r}
predict(model,
        newdata = data.frame(num_prior_arrests = 5))
```

Predicted time served = ______ months


## Question 6

Look at the R-squared value in the regression output.

```{r}
summary(model)
```

R-squared = ______

Interpret the R-squared in words.

Can we conclude that having more prior arrests **causes** a person to serve more time? Why or why not?

**Answer:**

