---
title: "Linear Regression In-Class Exercises - Solutions"
author: "Maria Cuellar"
date: "2026-10-06"
output: html_document
---



```{r setup, message=FALSE}
library(tidyverse)

dat <- read_csv("/Users/mariacuellar/Github/crim_data_analysis/data/riverton-crime.csv")
dat
```

# Linear Regression

Use riverton-crime.csv.

We will examine the relationship between the number of prior arrests and time served.

## Question 1

Create a scatterplot with `num_prior_arrests` on the x-axis and `time_served_months` on the y-axis.

```{r}
dat %>%
  ggplot(aes(x = num_prior_arrests, y = time_served_months)) +
  geom_point()
```

Describe the relationship. Is it positive or negative? Does it look approximately linear?

**Answer:** There is a weak positive relationship: people with more prior arrests tend to have somewhat higher time served. The relationship is approximately linear, but there is a lot of variation around the general trend.


## Question 2

Fit a linear regression where you regress time served onto number of prior arrests.

```{r}
model <- lm(time_served_months ~ num_prior_arrests, data = dat)

summary(model)
```

Which variable is the outcome/dependent variable?

**Answer:** `time_served_months`

Which variable is the explanatory/independent variable?

**Answer:** `num_prior_arrests`


## Question 3

Use the regression output to write the regression equation.

Predicted time served = 31.82 + 0.86 × number of prior arrests

Interpret the slope in words.

**Answer:** For each additional prior arrest, predicted time served increases by about 0.86 months, on average.


## Question 4

Add the regression line to the scatterplot.

```{r}
dat %>%
  ggplot(aes(x = num_prior_arrests, y = time_served_months)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

Does the regression line seem like a reasonable summary of the relationship?

**Answer:** It captures the slight positive trend in the data, but the observations are widely scattered around the line. So the line is a reasonable summary of the average trend, but number of prior arrests alone does not explain very much of the variation in time served.


## Question 5

Use the model to predict time served for a person with 5 prior arrests.

```{r}
predict(model,
        newdata = data.frame(num_prior_arrests = 5))
```

Predicted time served = **36.14 months**


## Question 6

Look at the R-squared value in the regression output.

```{r}
summary(model)
```

R-squared = **0.014**

Interpret the R-squared in words.

**Answer:** About 1.4% of the variation in time served is explained by its linear relationship with number of prior arrests. This is a very small proportion, so prior arrests alone are not a strong predictor of time served in this dataset.

Can we conclude that having more prior arrests **causes** a person to serve more time? Why or why not?

**Answer:** No. This regression shows an association between prior arrests and time served, not a causal effect. Other factors, such as offense type or seriousness, could be related to both prior arrests and time served. Establishing causation would require an appropriate research design, not just a regression.





## Question 6


Now let's look at the income_2.csv dataset. 
- Load the data
- Draw the scatterplot (without a line of best fit) (2 points)
- Describe the scatterplot, calculate correlation if appropriate (5 points)
- Fit a linear model and report the estimated i) slope, y-intercept, and iii) p-value for the slope. (5 points)
- Test the assumptions of linear regression and assess the goodness of fit. (5 points)

```{r}

income_dat <- read_csv("/Users/mariacuellar/Github/crim_data_analysis/data/income_2.csv")

income_dat %>% summarise(cor=cor(income,happiness))

income_dat %>%
  ggplot(aes(x = income, y = happiness)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

model <- lm(happiness~income, data=income_dat)

summary(model)

plot(model)

```

