---
title: "Linear Regression In-Class Exercises"
author: "Maria Cuellar"
date: "2026-10-05"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)

dat <- read_csv("/Users/mariacuellar/Github/crim_data_analysis/data/riverton-crime.csv")
```

# Linear Regression

We will examine the relationship between the number of prior arrests and time served.

## Question 1

Create a scatterplot with `num_prior_arrests` on the x-axis and `time_served_months` on the y-axis.

```{r}


```

Describe the relationship. Is it positive or negative? Does it look approximately linear?

**Answer:**



## Question 2

Fit a linear regression where you regress time served onto number of prior arrests.

```{r}

```

Which variable is the outcome/dependent variable?

**Answer:**

Which variable is the explanatory/independent variable?

**Answer:**


## Question 3

Use the regression output to write the regression equation.

Predicted time served = ______ + ______ × number of prior arrests

Interpret the slope in words.

**Answer:**


## Question 4

Add the regression line to the scatterplot.

```{r}


```

Does the regression line seem like a reasonable summary of the relationship?

**Answer:**


## Question 5

Use the model to predict time served for a person with 5 prior arrests.

```{r}
```

Predicted time served = ______ months


## Question 6

Look at the R-squared value in the regression output.

```{r}

```

R-squared = ______

Interpret the R-squared in words.

Can we conclude that having more prior arrests **causes** a person to serve more time? Why or why not?

**Answer:**






## Question 7


Now let's look at the income_2.csv dataset. 
a. Load the data
b. Draw the scatterplot (without a line of best fit) (2 points)
c. Describe the scatterplot, calculate correlation if appropriate (5 points)
d. Fit a linear model and report the estimated i) slope, y-intercept, and iii) p-value for the slope. (5 points)
e. Test the assumptions of linear regression and assess the goodness of fit. (5 points)





