---
title: "Exam 1 Questions"
author: "Maria Cuellar"
date: "2024-09-30"
output:
  html_document:
  self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Note: For this exam, don't worry about including the correct labels and titles for the plots.*

```{r preamble, echo=FALSE, message=FALSE}
library(tidyverse)
```

### 1. Loading data (5 points)

**Question:**

Load the data called riverton-crime.csv using the proper R command and file path.

**Code (5 points):** 

```{r load-data, message=FALSE}
2+2
```



### 2. Variable types (5 points)

**Question:**

Read the codebook. What types of (stat) variables are these? Variables: id, gender, offense_type, num_prior_arrests, and time_served_months.

**Text answer (5 points):**

- id: 
- gender: 
- offense_type: 
- num_prior_arrests: 
- time_served_months: 




### 3. Data dimensions (5 points)

**Question:**

How many observations and variables are there in the data? 


**Code (3 points):**

```{r}
2+2
```

**Text answer (2 points):**

Your answer.






### 4. Quantitative EDA (5 points)

**Question:**

What percentage of the sample is male?



**Code (3 points):**

```{r var-types}
2+2
```



**Text answer (2 points):**

Your answer.






### 5. Visual EDA (5 points)

**Question:**

Make a histogram for number of prior arrests and look for an appropriate number of bins. Describe this histogram.



**Code (3 points):**

```{r, fig.width=3, fig.height=3}
2+2
```



**Text answer (2 points):**

Your answer.





### 6. Quantitative EDA (5 points) 

**Question:**

Find statistics to describe the centrality and spread of the variable number of prior arrests. Are these numbers surprising?



**Code (3 points):**

```{r}
2+2
```


**Text answer (2 points):**

Your answer.





### 7. Visual EDA (5 points)

**Question:**

Now split up the histogram by gender. Does it look like males tend to have more or fewer prior arrests than females?



**Code (3 points):**

```{r, fig.width=5, fig.height=3}
2+2
```


**Text answer (2 points):**

Your answer.






### 8. Visual EDA (5 points) 


**Question:**

Make a barplot for offense_type and describe it.



**Code (3 points):**

```{r, fig.width=5, fig.height=3}
2+2
```

**Text answer (2 points):**

Your answer.






### 9. Visual EDA (5 points)

**Question:**

Now split the barplot by gender. Does it look like males commit more violent crimes than females?



**Code (3 points):**

```{r, fig.width=5, fig.height=3}
2+2
```

**Text answer (2 points):**

Your answer.






### 10. Visual EDA (5 points)

**Question:**

Make a histogram for time served. Split it up by offense type. Does it look like public order offenses have shorter or longer time served?



**Code (3 points):**

```{r, fig.width=5, fig.height=3}
2+2
```

**Text answer (2 points):**

Your answer.







### Extra credit. Quantitative EDA (5 points extra credit)

**Question:**

Now find a measure of centrality to determine whether your guess to #10 is correct. 

*Note: I have given you part of the code. You need to fill in the rest: Write the statistic you want to see in the empty round brackets. Next in the class, we will learn how to test these differences statistically.*

**Code (3 point):**

```{r}
2+2
```

**Text answer (2 point):**

Your answer.


