---
title: "Exercises 7 - questions"
author: "Maria Cuellar"
date: "2025-10-24"
output:
  html_document: 
  self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r preamble, echo=FALSE, message=FALSE}
library(tidyverse)
```


**Note**: The function `pnorm()` gives the cumulative probability for a **normal distribution** up to a certain value. That is, it tells you \( P(X \leq x) \). Use the problems below to practice computing and interpreting probabilities with the normal distribution.



## Problem 1: Basic Standard Normal Probability

Find the probability that a standard normal variable \( Z \) is less than 1.25.
What does this number represent?


## Problem 2: Two-Sided Probability

Compute the probability that Z lies between -1.5 and 1.5.


**Note**: The rest of the problems are about t-tests.

## Problem 3: t-test 

- What is a t-test for?
- What types of t-tests are there? 
- What are the assumptions for t-tests?


## Problem 4: Perform a one-sample t-test

A criminologist believes that the average daily number of arrests in a city is less than 25. She collects a random sample of 15 days and finds the data in problem1_arrests.csv.

- State the null and alternative hypotheses.
- Check assumptions for a one-sample t-test (random sample, approximate normality).
- Perform the t-test.
- Report and interpret the test statistic, degrees of freedom, and p-value.
- Write a short paragraph interpreting the results in context.




## Problem 5: Perform a two-sample t-test

Researchers compare the mean sentence length (in months) given by two judges. Random samples of sentences are recorded in problem2_judges.csv.

- State H0 and Ha.
- Check assumptions (independence, approximate normality, equal variances if needed).
- Conduct both:
  - a pooled (equal variances) t-test, and
  - a Welch (unequal variances) t-test.
- Compare results. Which test seems more appropriate?
- Interpret in context.



## Problem 6: Perform a Paired t-test

Police cadets are tested on physical fitness before and after a 10-week training program: see problem3_cadets.csv.

- Why is this a paired design?
- Conduct a paired t-test.
- Report the t statistic, df, and p-value.
- Interpret whether the training improved fitness.



## Problem 7: Inference with linear regression

Do inference on the slope coefficient for the regression you performed in Exercises 5, lm(price ~ sqft, data=dat), using philadelphia_house_prices.csv.


