---
title: "Exercises 6 - questions"
author: "Maria Cuellar"
date: "2025-10-13"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r preamble, echo=FALSE, message=FALSE}
library(tidyverse)
```

Note: Continue using philadelphia_house_prices.csv.
```{r loading-data}
dat <- read_csv("/Users/mariacuellar/Github/crim_data_analysis/data/philadelphia_house_prices.csv")
```


**1. Question** For two quantitative variables (x and y): Write down the requirements for each of these steps.

- Do visual EDA:
- Do quantitative EDA:
- Fit a linear model:
- Test assumptions of linear regression:
- Interpret coefficient of x to do inference:



**Instructions** Fit a simple linear regression for price regressed onto square footage. 

**2. Question** What is the null hypothesis for doing inference? What is the research question you could ask here?

**3. Question** What is the coefficient of determination? (The one called Multiple R-squared. The other one has a penalty for adding more covariates, so we won't need it here.) What does this mean?

**4. Question** Test the assumptions of the linear regression. 

a) Linear relationship between x and y:
b) Independence between observations:
c) Homoscedasticity of errors:
d) Normality of y for a given x: 


**5. Question** Should you interpret the coefficients? If so, go ahead and interpret the coefficients. If not, then give a satatement for why you should not interpret them.






