---
title: "Exercises 5 - questions"
author: "Maria Cuellar"
date: "2025-10-06"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r preamble, echo=FALSE, message=FALSE}
library(tidyverse)
```

Note: Use philadelphia_house_prices.csv.


**1. Question:** Do visual EDA on price, and do visual EDA on square footage. Describe what you see.

**2. Question:** Could there be a categorical variable that explains the strange features observed in price? Look for it in the data.

**3. Question** Do quantitative EDA on price and square footage, separately. Describe.

**4. Question** Do visual EDA on price and square footage, together. Describe what you see.

**5. Question** Do quantitative EDA on price and square footage, together. Describe.

**6. Question**  Repeat 3, but now add a color for the categorical variable you found in 2. Describe what you see.

**7. Question**  Fit a simple linear regression for price and square footage. What is your null hypothesis? Look at the summary. For now, don't interpret the coefficients and don't test for the assumptions.

