---
title: "Exercises 5 - solutions"
author: "Maria Cuellar"
date: "2025-10-06"
output:
  html_document: 
  self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r preamble, echo=FALSE, message=FALSE}
library(tidyverse)
```

```{r}
dat <- read_csv("/Users/mariacuellar/Github/crim_data_analysis/data/philadelphia_house_prices.csv")
dat
```



# 1. Do visual EDA on price, and do visual EDA on square footage. Describe what you see.

```{r, fig.width=3, fig.height=3}
dat %>% ggplot(aes(x=price)) + geom_histogram()
dat %>% ggplot(aes(x=sqft)) + geom_histogram()
```

# 2. Could there be a categorical variable that explains the strange features observed in price? Look for it in the data.

```{r}
dat %>% ggplot(aes(x=price, fill=neighborhood)) + geom_histogram()
```

# 3. Do quantitative EDA on price and square footage, separately. Describe.
```{r}
dat %>%
  summarize(
    mean_price = mean(price, na.rm = TRUE),
    sd_price   = sd(price, na.rm = TRUE),
    mean_sqft  = mean(sqft, na.rm = TRUE),
    sd_sqft    = sd(sqft, na.rm = TRUE)
  )
```

# 4. Do visual EDA on price and square footage, together. Describe what you see.
```{r}
dat %>% ggplot(aes(x=price, y=sqft)) + geom_point()
```

# 5. Do quantitative EDA on price and square footage, together. Describe.
```{r}
dat %>% summarize(correlation = cor(price, sqft))
```

# 6. Repeat 3, but now add a color for the categorical variable you found in 2. Describe what you see.
```{r}
dat %>% ggplot(aes(x=price, y=sqft, color=neighborhood)) + geom_point()
```

# 7. Fit a simple linear regression for price and square footage. What is your null hypothesis? Look at the summary. For now, don't interpret the coefficients and don't test for the assumptions.
```{r}
out <- lm(price ~ sqft, data=dat)
summary(out)
```

