---
title: "Exercises 7 - questions"
author: "Maria Cuellar"
date: "2025-10-24"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r preamble, echo=FALSE, message=FALSE}
library(tidyverse)
```

Note: Use community_center_crime.csv.

```{r}
dat <- read_csv("/Users/mariacuellar/Github/crim_data_analysis/data/community_center_crime.csv")
```

**2. Question:** You are analyzing data from 20 neighborhoods in a city. Ten of them have a community center, and ten do not. You want to test whether neighborhoods with community centers have a lower average property crime rate (per 1,000 residents) than those without community centers. 

Variables:

community_center: 1 = Yes, 0 = No

crime_rate: property crimes per 1,000 residents

Using a two-sample t-test, test whether the mean crime rate differs between neighborhoods with and without community centers.


State the hypotheses (both null and alternative).

Check assumptions (independence, approximate normality, similar variance).

Perform the t-test in R:

```{r}
t.test(crime_rate ~ community_center, data = dat, var.equal = TRUE)
```

Interpret the p-value in context.

Write one sentence about what the result means for crime prevention policy.


