---
title: "Exercises 8"
author: "Your name"
date: "2024-10-07"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

# Exercises 8

```{r preamble}
library(tidyverse)
```

### 1. Load the data called penguin_dives.csv. 

```{r load-data}
dat <- read_csv("data/penguin_dives.csv")
```

### 2. Draw a scatterplot of x=DiveHeartRate, y=Duration. What do you see there?

```{r}

dat %>% ggplot(aes(x=DiveHeartRate, y=Duration)) + geom_point()

```

### 3. Fit a linear model of DiveHeartRate vs. Duration and summarize it.

```{r linear-model}

cor(dat$DiveHeartRate, dat$Duration)

out <- lm(Duration ~ DiveHeartRate, data=dat)
summary(out)

```

### 4. Draw the diagnostic plots for the model. What do you see regarding the assumptions? 

```{r diagnostic-plots}

plot(out)

```


### 5. Try out different transformations, ^2, ^.5, log.

```{r transformations}

dat %>% ggplot(aes(x=DiveHeartRate, y=(Duration)^2)) + geom_point() + stat_smooth(method = "lm", se = TRUE)
dat %>% ggplot(aes(x=DiveHeartRate, y=(Duration)^.5)) + geom_point() + stat_smooth(method = "lm", se = TRUE)
dat %>% ggplot(aes(x=DiveHeartRate, y=log(Duration))) + geom_point() + stat_smooth(method = "lm", se = TRUE)




```

### 6. Now fit the linear model again with the transformed data. Are the diagnostics better?

```{r lm-transformed}

out_transformed <- lm(log(Duration) ~ DiveHeartRate, data=dat)
plot(out)
plot(out_transformed)
```


### 7. Explore the outliers.
```{r outliers}


```

