---
title: "Exercises 4 - solutions"
author: "Maria Cuellar"
date: "2025-09-22"
output:
  html_document:
  self_contained: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r preamble}
library(tidyverse)
```

# 1. A summary statistic, mean of disp

```{r}
mtcars %>% summarize(mean=mean(disp))
```

# 2. A histogram of disp
```{r, fig.width=5, fig.height=3}
mtcars %>% ggplot(aes(x=disp)) + geom_histogram()
```


# 3. A short paragraph explaining what you found

This histogram is difficult to describe because it does not seem to be part of a clear, continuous distribution. Even if we play around with the number of bins (using bins=10, or bins=50). You can give the modes (one, or three), say it's skewed, and that there are no clear outliers. There is no correct answer: you just need to describe what you see. 

This is exploratory data analysis, and the goal is to generate hypotheses that we will test statistically.

# Submit the html file on canvas.

*Note that the "self_contained: true" section makes it so that you produce an HTML file that contains all the graphics and everything needed to see your results. And that's what you'll submit on canvas.*