---
title: "Quiz 2 Practice - Questions"
author: "Your name"
date: "2025-10-15"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



> Use the pretrial_df dataset. 

```{r load-data, echo=FALSE, message=FALSE}
library(tidyverse)
dat <- read_csv("/Users/mariacuellar/Github/crim_data_analysis/data/pretrial_df.csv")
```


---


### 1. Visual EDA (5 points)

> Do visual EDA on the variables bail_amount and days_until_trial, jointly.

**Your code answer (5 points):**

```{r fig.width=4, fig.height=3}
dat %>% ggplot(aes(x=bail_amount, y=days_until_trial)) + geom_point()
```

**Your text answer (5 points):**

The scatterplot shows that the two variables have a positive association, a linear form (except for the truncation at 0), the relationship is not that strong, and there are some outliers, especially in high days until trial. 




---


### 2. Quantitative EDA (5 points)

> Do quantitative EDA on the variables bail_amount and days_until_trial, jointly.

**Your code answer (5 points):**

```{r}
dat %>% summarize(correlation = cor(bail_amount, days_until_trial))

dat %>% filter(days_until_trial < 200) %>% summarize(correlation = cor(bail_amount, days_until_trial))
```


**Your text answer (5 points):**

After checking that we are working with two quantitative variables, the relationship between the two variables is straight enough, we fit a correlation. There is at least an outlier in days_until_trial, so we provide the correlation with and without the outliers. The correlation went from 48.5% to 48.2%, so the outlier was not making a big difference. 



---





### 3. Linear model fit (5 points)

> Fit a linear model that regresses bail_amount onto days_until_trial. Are the assumptions of linear regression met? Are there any influential outliers?

**Your code answer (5 points):**

```{r, fig.width=8, fig.height=6}
out <- lm(bail_amount~days_until_trial, data=dat)
par(mfrow = c(2,2))
plot(out)
par(mfrow = c(1,1))
```



**Your text answer (5 points):**

- Linearity: Not really. The truncation at 0 makes it hard to tell. There is clearly a trend that is not being captured by the model.
- Independence of observations: We have no reason to think that there are hidden dependencies in the data.
- Normality of errors: The points are mostly on the diagonal in the q-q plot, except for the tails. So, normality is mostly met.
- Homoscedasticity: This is mostly fine except for the high tail. We don't know if there really is lower variance, or just fewer data points at the extreme.
- Influential outliers: None according to the Cook's distance.


---





### 4. Slope coefficient interpretation (5 points)

> Should you interpret the slope coefficient of the linear model?

**Your text answer (5 points):**

No because the model assumptions are not met. A linear regression is not the right model to fit here.


---




