---
title: "Quiz 2  - Questions"
author: "Your name"
date: "2025-10-15"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



> Use the pretrial_df dataset. 

```{r load-data, echo=FALSE, message=FALSE}
library(tidyverse)
2+2
```


---


### 1. Visual EDA (5 points)

> Do visual EDA on the variables bail_amount and days_until_trial, jointly.

**Your code answer (5 points):**

```{r fig.width=4, fig.height=3}
2+2
```

**Your text answer (5 points):**





---


### 2. Quantitative EDA (5 points)

> Do quantitative EDA on the variables bail_amount and days_until_trial, jointly.

**Your code answer (5 points):**

```{r}
2+2
```


**Your text answer (5 points):**




---





### 3. Linear model fit (5 points)

> Fit a linear model that regresses bail_amount onto days_until_trial. Are the assumptions of linear regression met? Are there any influential outliers?

**Your code answer (5 points):**

```{r, fig.width=8, fig.height=6}
2+2
```



**Your text answer (5 points):**

-
-
-
-
- 

---





### 4. Slope coefficient interpretation (5 points)

> Should you interpret the slope coefficient of the linear model?

**Your text answer (5 points):**



---




