---
title: "Practice for Exam 1"
author: "Your name"
date: "2024-09-30"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Practice for Exam 1

```{r preamble}
library(tidyverse)
```

### 1. Load the data called police_use_of_force.

```{r load-data}
dat <- read_csv("police_use_of_force.csv")

```
Answer: 


### 2. What variables are included in the dataset?

```{r what-variables}
names(dat)
```

Answer:


### 3. What type of stat variables are ForceType, Race, and EventAge and what values can they take? Comment on anything unusual you find.

```{r age2-type}
levels(as.factor(dat$ForceType))
levels(as.factor(dat$Race))

range(dat$EventAge, na.rm = TRUE) # looks like there's an erroneous value
dat$EventAge <- ifelse(dat$EventAge<0, NA, dat$EventAge) # this removes anything below 0
range(dat$EventAge, na.rm = TRUE) # looks better now, from 0-82.
```

Answer: ForceType and Race are Categorical and EventAge is quantitative. 

ForceType:
[1] "Baton"                       "Bodily Force"                "Chemical Irritant"          
 [4] "Firearm"                     "Gun Point Display"           "Improvised Weapon"          
 [7] "Less Lethal"                 "Less Lethal Projectile"      "Maximal Restraint Technique"
[10] "Police K9 Bite"              "Taser"  

Race:
[1] "Asian"              "Black"              "Native American"    "not recorded"       "Other / Mixed Race"
[6] "Pacific Islander"   "Unknown"            "White"             

EventAge: Ranges from 0-82 when erroneous data is removed.


### 4. Do visual EDA for ForceType and Race. Do you find anything interesting or unusual?

```{r ForceType-visual-EDA}
dat %>%
  ggplot(aes(ForceType)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1)) # this line makes the labels vertical

dat %>%
  ggplot(aes(Race)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1)) # this line makes the labels vertical
```

Answer: 


### 5. Do quantitative EDA for ForceType and Race. Which do you think is more helpful, a plot or a table?

```{r ForceType-quant-EDA}
dat %>%
  count(ForceType) %>%
  mutate(prop = prop.table(n))

dat %>%
  count(Race) %>%
  mutate(prop = prop.table(n))
```

Answer: 


### 6. Do visual EDA on EventAge. What can you learn from this?

```{r EventAge-visual-EDA}
hist(dat$EventAge)

dat %>% 
  ggplot(aes(EventAge)) + 
  geom_histogram()

```

Answer:


### 7. Do quantitative EDA for EventAge.

```{r EventAge-quant-EDA}
dat %>%
  filter(!is.na(EventAge)) %>% 
  summarize(median = median(EventAge), 
            IQR = IQR(EventAge),
            n = length(EventAge))
```

Answer: 

