---
title: "Exercises 5"
author: "Maria Cuellar"
date: "2024-09-20"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
# echo=FALSE hides the code in the knitted file, include=FALSE doesn't include that chunk.
```

# Exercises 5

```{r preamble}
# install.packages("tidyverse")
library(tidyverse)
```

### 1. Load the data called domestic violence

```{r load-data}
dat <- read_csv("data/domestic_violence.csv") # make sure data is in the right folder
```

Answer: Loaded the data.

### 2. What type of stat variable is Employment?

```{r Employment-type}
dat$Employment # base R version

dat %>% select(Employment) # tidyverse version, pick out the variable Employment to look at it
```

Answer: Employment is a categorical variable.


### 3. What categories does it have?

```{r Employment-categories}
levels(as.factor(dat$Employment)) # base R version

dat <- dat %>% mutate(Employment=as.factor(Employment)) # tidyverse version, make Employment a factor

levels(dat$Employment) # read the levels of the factor
```

Answer: Employed, semi employed, and unemployed.


### 4. Do quantitative EDA for Employment.

```{r Employment-quant-EDA}
table(dat$Employment)

dat %>% 
  count(Employment) %>%
  mutate(prop = prop.table(n)) # make a table of counts, this one includes proportions.
```

Answer: I made a table of Employment, and I can see the three categories and their respective counts, as well as their proportions.

### 5. Do visual EDA for Employment.

```{r Employment-visual-EDA}
barplot(table(dat$Employment)) # base R


dat %>%
  ggplot(aes(x=Employment)) +
  geom_bar() +
  theme_minimal()

```

Answer: Made a barplot. 

### 6. What kind of variables are `Marital status` (and why does it have single quotes around it) and Violence?

Categorical, and `Marital status` has quotes because it has a space in the name.


### 7. Make a contingency table of both Marital status and Violence
```{r contingency-table}
addmargins(table(dat$`Marital status`, dat$Violence)) # use base-R to make this. Is there a better way?
```

Answer: Made a contingency table of the two categorical variables, with margins.

