---
title: "Exam 1"
author: "Your name"
date: "2024-09-30"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exam 1

```{r preamble}

```

### 1. (5 points) Load the data called dat.nsduh.small.csv. Take a look at the data.

```{r load-data}

```

Answer:


### 2. (5 points) Read the codebook. What do these variables represent: mjage, iralcage, sexatract, speakengl, and irsex?

Answer: 


### 3. (5 points) What type of stat variables are the variables from #2?

```{r var-types}

```

Answer: 


### 4. (5 points) Do visual EDA for mjage and iralcage, separately. (Hint: To compare the plots, make the axes limits, x and y, be the same for both plots. Also, try out different numbers of bins to see how they change.)

```{r visual-EDA-quantitative}

```

Answer: 


### 5. (5 points) Compare the two plots. How are they different and how are they similar? What theory do you have to explain these differences?

Answer: 


### 6. (5 points) Do visual EDA for irsex and sexatract. What do you learn from this?

```{r visual-EDA-categorical}

# Note: This line makes the levels of the factors take on the values from the codebook. Run it before you make your plots to have clearer labels.
dat <- dat %>%
  mutate_at(c('irsex', 'sexatract', 'speakengl'), as.factor) %>%
  mutate(irsex = recode(irsex,         "1"="Male", "2"="Female"),
         sexatract = recode(sexatract, "1"="I am only attracted to opposite sex",
                                       "2"="I am mostly attracted to opposite sex",
                                       "3"="I am equally attracted to males and females",
                                       "4"="I am mostly attracted to same sex",
                                       "5"="I am only attracted to same sex",
                                       "6"="I am not sure",
                                       "99"="Legitimate skip"),
         speakengl = recode(speakengl, "1"="Very well",
                                       "2"="Well",
                                       "3"="Not well",
                                       "4"="Not at all"))


```

Answer: 


### 7. (5 points) Do quantitative EDA for irsex, sexatract, and speakengl. What do you learn from this that you didn't know from the visual EDA?

```{r quantitative-EDA-categorical}

```

Answer: 


### 8. (5 points) Copy your line from the visual EDA for mjage. What happens when you use this code "aes(x=mjage, fill=sexatract)" for the aesethetics in ggplot? Do you find out something interesting?

```{r adding-fill}

```

Answer: 
